### 2016-06-29 publications/Journal/4a-allocation.tex
Phold and PHOLD consistently

   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_BF.pdf}
   \caption{PholdTree model breadth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_bfs}
   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_DF.pdf}
   \caption{PholdTree model depth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_dfs}
\end{figure}

@@

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4BFS.eps}
    \caption{Visualization of a PholdTree simulation with breadth first allocation and 4 kernels.}
    \label{fig:pholdtree_visualize_parBFS}
\end{figure}
\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4DFS.eps}
    \caption{Visualization of a PholdTree simulation with depth first allocation and 4 kernels.}
    \label{fig:pholdtree_visualize_parDFS}
\end{figure}

@@

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreeallochighp.eps}
    \caption{PholdTree model performance using different allocation schemes.}
    \label{fig:PholdTree_plot_alloc_high}
\end{figure}

>>>>

   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_BF.pdf}
   \caption{PHOLDTree model breadth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_bfs}
   \includegraphics[width=\columnwidth]{fig/pholdtree_alloc_DF.pdf}
   \caption{PHOLDTree model depth first allocation with 4 kernels.}
   \label{fig:PholdTree_model_dfs}
\end{figure}

@@

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4BFS.eps}
    \caption{Visualization of a PHOLDTree simulation with breadth first allocation and 4 kernels.}
    \label{fig:pholdtree_visualize_parBFS}
\end{figure}
\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreed1n3t5000c4DFS.eps}
    \caption{Visualization of a PHOLDTree simulation with depth first allocation and 4 kernels.}
    \label{fig:pholdtree_visualize_parDFS}
\end{figure}

@@

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/pholdtreeallochighp.eps}
    \caption{PHOLDTree model performance using different allocation schemes.}
    \label{fig:PholdTree_plot_alloc_high}
\end{figure}

### 2016-06-29 publications/Journal/parallel-performance.tex
Phold and PHOLD consistently

\subsection{Parallel Simulation}
We now continue by describing our parallel simulation performance for different synchronization protocols, and compared to adevs.
This is done for our three benchmark models.

@@

Figure~\ref{fig:Queue_plot_strong} shows the speedup compared to dxex sequential for a fixed problem size.
As the amount of kernels increases, the optimistic kernel quickly becomes the worst choice.
The difference between dxex conservative and adevs becoming smaller.
The same effect can be seen for weak scaling in Figure~\ref{fig:Queue_plot_weak}.

The speedup of adevs is always in comparison with the runtime of the corresponding dxex sequential benchmark.
\begin{figure}
	\center

	\includegraphics[width=\columnwidth]{fig/queue_fixed_weak_speedup.eps}
	\caption{Queue model weak scaling speedup compared to dxex sequential.}
	\label{fig:Queue_plot_weak}

	\includegraphics[width=\columnwidth]{fig/queue_fixed_strong_speedup.eps}
	\caption{Queue model strong scaling speedup compared to dxex sequential.}
	\label{fig:Queue_plot_strong}

\end{figure}

\subsubsection{Interconnect}\label{subsec:parallelinterconnect}
In the Interconnect model, we determine how broadcast communication is supported across multiple nodes.
The number of models is now kept constant at eight.
Results are shown in Figure~\ref{fig:interconnect_benchmark_parallel}.
When the number of nodes increases, performance decreases due to increasing contention in conservative simulation and an increasing number of of rollbacks in optimistic simulation.
All models depend on each other and have no computational load whatsoever, negating any possible performance gain by executing the simulation in parallel.

@@

\subsubsection{Phold}
In the Phold model, we first investigate the influence of the percentage of remote events on the speedup. A remote event in this context is an event that is sent from a model on one kernel to a model on another simulation kernel.
When remote events are rare, optimistic synchronization rarely has to roll back, thus increasing performance.
With more common remote events, however, optimistic synchronization quickly slows down due to frequent rollbacks.
Conservative synchronization, on the other hand, is mostly unconcerned with the number of remote events: the mere fact that a remote event can happen, causes it to block and wait.
Even though a single synchronization protocol is always ideal in this case, it already shows that different synchronization protocols respond differently to a changing model.

Adevs is significantly slower during conservative synchronization.
Analysis of profiling callgraphs shows that exception handling in adevs is the main cause. 
To keep the models equivalent, the adevs version does not provide the \{begin,end\}Lookahead methods, which accounts for the exception handling.
These functions require the user to implement a state saving in contrast to PythonPDEVS and dxex's optimistic kernels which handle this inside the kernel.
We feel this would lead to an unfair comparison as we would like to keep the models synchronization-agnostic across all benchmarks.

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/phold_remotes.eps}
    \caption{Phold benchmark results for parallel simulation using four kernels, four atomics per node, with varying percentage of remote events.}
\end{figure}
\begin{figure}
	\center
	\includegraphics[width=\columnwidth]{fig/phold_priority.eps}
	\caption{Phold benchmark results for parallel simulation using four kernels, with varying amount of high-priority events.}
	\label{fig:phold_priority}
\end{figure}

We slightly modified the Phold benchmark, to include high-priority events.
Contrary to normal events, high-priority events happen almost instantaneously, restricting lookahead to a very small value.
Even when normal events occur most often, conservative synchronization always blocks until it can make guarantees.
Optimistic synchronization, however, simply goes forward in simulation time and rolls back when these high-priority events happen.
This situation closely mimics the case made in the comparison between both synchronization algorithms by~\cite{FujimotoBook}.

>>>>

\subsection{Parallel Simulation}
We now continue by describing our parallel simulation performance for different synchronization protocols, and compared to adevs.
The speedup of adevs is computed with the corresponding dxex sequential benchmark.
This was done to take into account the performance difference observed in sequential simulation.

@@

Figure~\ref{fig:Queue_plot_strong} shows the speedup compared to sequential simulation for a fixed problem size.
%TODO this is actually quite a strange choice...
As the number of kernels increases, the optimistic kernel quickly becomes the worst choice.
This is mainly caused by the pipeline structure of the model: the last models in the queue only respond to incoming messages and therefore have to be rolled back frequently.
The difference between dxex conservative and adevs becoms smaller when more and more cores are used.
The same effect can be seen for weak scaling in Figure~\ref{fig:Queue_plot_weak}.

\begin{figure}
	\center
	\includegraphics[width=\columnwidth]{fig/queue_fixed_weak_speedup.eps}
	\caption{Queue model weak scaling speedup compared to dxex sequential.}
	\label{fig:Queue_plot_weak}
\end{figure}

\begin{figure}
	\center
	\includegraphics[width=\columnwidth]{fig/queue_fixed_strong_speedup.eps}
	\caption{Queue model strong scaling speedup compared to dxex sequential.}
	\label{fig:Queue_plot_strong}
\end{figure}

\subsubsection{Interconnect}\label{subsec:parallelinterconnect}
In the Interconnect model, we determine how broadcast communication is supported across multiple nodes.
The number of models is now kept constant at eight.
Results are shown in Figure~\ref{fig:interconnect_benchmark_parallel}.
When the number of nodes increases, performance decreases due to increasing contention in conservative simulation and the increasing number of rollbacks in optimistic simulation.
All models depend on each other and have no computational load whatsoever, negating any possible performance gain by executing the simulation in parallel.

@@

\subsubsection{PHOLD}
In the PHOLD model, we first investigate the influence of the percentage of remote events on the speedup.
A remote event in this context is an event that is sent from a model on one kernel to a model on another simulation kernel.
When remote events are rare, optimistic synchronization rarely has to roll back, thus increasing performance.
With more common remote events, however, optimistic synchronization quickly slows down due to frequent rollbacks.
Conservative synchronization, on the other hand, is mostly unconcerned with the number of remote events: the mere fact that a remote event can happen, causes it to block and wait.
Even though a single synchronization protocol is always ideal in this case, it already shows that different synchronization protocols respond differently to a changing model.

Adevs is significantly slower during conservative synchronization.
Analysis of profiling callgraphs shows that exception handling in adevs is the main cause. 
To keep the models equivalent, the adevs version does not provide the \{begin,end\}Lookahead methods, which accounts for the exception handling.
These functions require the user to implement a state saving method.
But in contrast to PythonPDEVS and dxex, which handle this inside the kernel, users need to manually define this.
We feel this would lead to an unfair comparison as we would like to keep the models agnostic of the underlying protocols across all benchmarks.

\begin{figure}
    \center
    \includegraphics[width=\columnwidth]{fig/phold_remotes.eps}
    \caption{PHOLD benchmark results for parallel simulation using four kernels, four atomics per node, with varying percentage of remote events.}
\end{figure}
\begin{figure}
	\center
	\includegraphics[width=\columnwidth]{fig/phold_priority.eps}
	\caption{PHOLD benchmark results for parallel simulation using four kernels, with varying amount of high-priority events.}
	\label{fig:phold_priority}
\end{figure}

We slightly modified the PHOLD benchmark, to include high-priority events.
Contrary to normal events, high-priority events happen almost instantaneously, restricting lookahead to a very small value.
Even when normal events occur most often, conservative synchronization always blocks until it can make guarantees.
Optimistic synchronization, however, simply goes forward in simulation time and rolls back when these high-priority events happen.
This situation closely mimics the case made in the comparison between both synchronization algorithms by~\cite{FujimotoBook}.
